---
title: "Homework #5"
author: L. Harvey 
embed: TRUE 
echo: FALSE
format: html
editor: visual
---

# Homework Assignment #5

## Week 5 (Notes)

Get familiar with the following distributions/functions in R:

```{r}
?distributions #This command gives us a list of distributions in the R Studio package, including details on different distibution types and a breakdown of their arguments. 
```

Example Code:

```{r}
input <- 350
output <- pbinom(input, size = 1000, prob = 0.345623)
qbinom(output, size = 1000, prob = 0.345623)
```

Background Code:

```{r}
library(tidyverse)
theme_set(theme_light())
```

## Exercise 5.2.1

```{r}
# For testing Equation 5.1 with dbinom: 
dbinom(x = 1, size = 1, prob = 2/3)
dbinom(x = 0, size = 1, prob = 2/3)
dbinom(x = 4, size = 1, prob = 2/3)

# For testing Equation 5.2 with qbinom: 
qbinom(-1, 1, 2/3)
qbinom(0, 1, 2/3)
qbinom(1, 1, 2/3)
qbinom(5, 1, 2/3)

# Use rbinom to generate 100 samples, then graph those results: 
hwsim <- rbinom(n = 100, size = 1, prob = 2/3)
hwframe <- data.frame(hwsim)
       
ggplot(hwframe, aes(hwsim)) +
  geom_bar()
```

## Exercise 5.3.1

If we flip a fair coin four times, there are five possible outcomes (where "Heads" = "H" and "Tails" = "T"), but there are sixteen EVENTS possible in this sample space:

0 Heads:

-   T, T, T, T

1 Heads:

-   T, T, T, H

-   T, T, H, T

-   T, H, T, T

-   H, T, T, T

2 Heads:

-   T, T, H, H

-   T, H, H, T

-   H, H, T, T

-   H, T, H, T

-   T, H, T, H

-   H, T, T, H

3 Heads:

-   T, H, H, H

-   H, T, H, H

-   H, H, T, H

-   H, H, H, T

4 Heads:

-   H, H, H, H

## Exercise 5.3.2

```{r}
# How many possible ways are there to get X = 0, X = 1, X = 2, X = 3, and X = 4?

# To find X = 0 
dbinom(0, 4, 0.5)
# To find X = 1
dbinom(1, 4, 0.5)
# To find X = 2
dbinom(2, 4, 0.5)
# To find X = 3
dbinom(3, 4, 0.5)
# To find X = 4
dbinom(4, 4, 0.5)
choose()
```

The probability that X = 2 is 37.5%, which we can verify with the code below:

```{r}
# Verifying that the probability of X = 2 is 37.5%
dbinom(2, 4, 0.5)

# It worked!!!! Now let's use the right R function to calculate the probability of X being less than 1: 
pbinom(1, 4, 0.5)
```

## Exercise 5.3.3

```{r}
# Use rbinom to generate 1000 samples from the Binomial distribution with size = 5 and probability = 2/3

# First, create values for this sample: 
hwsampler <- rbinom(1000, size = 5, prob = 2/3)

# Next, create the frame/dresser: 
big_boy_dresser <- data.frame(hwsampler)

# Now we can create a ggplot to show the results: 
ggplot(big_boy_dresser, 
       aes(x = hwsampler)) + 
  geom_bar()

# Lastly, let's estimate the probability that X is an even number, using the mean() function on the 1000 values in our hwsampler dataset. 
likelihood_even <- mean(hwsampler %% 2 == 0) 
likelihood_even
```

## Exercise 5.3.5

Think back to the shenanigans of Sections 5.3.1 and 5.3.2.

```{r}
 # Find the probability that X = 2 when probability = 1/3 and n = 4. 
dbinom(2, 4, 1/3)
```

## Exercise 5.3.7

```{r}
# If probability = 2/3 and n = 15, use dbinom() and pinom() to verify the results of equations 5.3 and 5.4
# (Also note: I was super confused by what we were supposed to do with this exercise. How do we "verify" an equation other than solving it with dbinom or pbinom?)

# Using dbinom) to verify Equation 5.3:
dbinom(x = seq(0, 15), 15, 2/3)

# Using pbinom) to verify Equation 5.4:
pbinom(q = seq(0, 15), size = 15, prob = 2/3)
```

## Exercise 5.4.1

```{r}
# Steve Code
library(tidyverse)
grid <- tibble(prob = seq(0, 1, by = 0.001))
grid$like <- dbinom(21, 47, grid$prob)

# Homework Code 
grid <- grid |> 
  mutate(clike_raw = cumsum(like), 
         clike_normalized = clike_raw / sum(like)
         )

grid |> filter (prob == 0.588)

# After normalizing clike, we find a probability of 97.5% that "prob" is equal to or greater than .588. 
```
